Mergeable Types

### 2017-02-28 paper/implementation.tex

\begin{wrapfigure}{L}{.5\textwidth}
	\begin{center}
	\includegraphics[scale=0.8]{Figures/partitions}
	\end{center}
	\caption{Network partitions may leave branches $b_1$ and $b_4$ in one
		partition (red background), and branches $b_2$ and $b_3$ in the other
		(yellow background). Since $b_2$ and $b_3$ cannot merge unless $b_1$ and
		$b_4$ merge, they cannot communicate among themselves even though they are in
		the same network partition.}
	\label{fig:partitions}
\end{wrapfigure}

Concurrent operations in Irmin are tracked using the notion of
branches, allowing the programmer to explicitly merge branches on
demand. \name's concurrency and distribution model is realized in
terms of branches; each replica operates on its own branch, and thus
is isolated from the actions on other replicas. In the presence of
network partitions, the nodes in one partition may not receive updates
from the other partition, preventing merges from happening across a
partition. This can be a problem if a pair of branches in one
partition (the current partition) have multiple LCAs in the other
partition (the remote partition), and hence cannot
merge. Fig.~\ref{fig:partitions} illustrate this problem for the
branching structure of Fig.~\ref{fig:external-lcas}.

@@

Importantly, Irmin supports user-defined three-way merge functions for
reconciling concurrent operations. While Irmin's merge functions are defined
over objects on Irmin's content-addressable heap, \name's merge functions are
defined over OCaml types. We address this representational mis-match with the
help of OCaml's PPX metaprogramming support~\cite{ppx} to derive bi-directional
transformations between objects on OCaml and Irmin heaps. We also derive the
various serialization functions required by Irmin. This address the representation
mismatch between the application and storage layers. Synchronization between
replicas is performed using Git's notion of pushing and pulling updates from
remotes over the Git transfer protocol~\cite{git-tp}. As a result, we reap the
benefits of efficient delta-transfer (only missing objects are transferred
between replicas), compression and end-to-end encrypted communication between
the replicas.

>>>>

\begin{wrapfigure}{L}{.5\textwidth}
	\begin{center}
	\includegraphics[scale=0.8]{Figures/partitions}
	\end{center}
  \caption{Network partitions may leave branches $b_1$ and $b_4$ in
  one partition (red background), and branches $b_2$ and $b_3$ in the
  other (yellow background). Access to full history on both sides of
  the partition lets both sides make progress.}
	\label{fig:partitions}
\end{wrapfigure}

Concurrent operations in Irmin are tracked using the notion of
branches, allowing the programmer to explicitly merge branches on
demand. \name's concurrency and distribution model is also realized in
terms of branches; each replica operates on its own branch, and thus
is isolated from the actions on other replicas. In the presence of
network partitions, the nodes in one partition may not receive updates
from the other partition, preventing merges from happening across a
partition. This can be a problem if a pair of branches in one
partition (the current partition) have multiple LCAs in the other
partition (the remote partition), and hence cannot
merge. Fig.~\ref{fig:partitions} illustrate this problem for the
branching structure of Fig.~\ref{fig:external-lcas}.

@@

Importantly, Irmin supports user-defined three-way merge functions to
reconcile concurrent operations.  While Irmin's merge functions are defined
over objects on Irmin's content-addressable heap, \name's merge functions are
defined over OCaml types. We address this representational mis-match with the
help of OCaml's PPX metaprogramming support~\cite{ppx} to derive bi-directional
transformations between objects on OCaml and Irmin heaps. We also derive the
various serialization functions required by Irmin. This address the representation
mismatch between the application and storage layers. Synchronization between
replicas is performed using Git's notion of pushing and pulling updates from
remotes over the Git transfer protocol~\cite{git-tp}. As a result, we reap the
benefits of efficient delta-transfer (only missing objects are transferred
between replicas), compression and end-to-end encrypted communication between
the replicas.
